---
title: "Lab 5"
author: "Deanna Elliott"
date: "10/26/2021"
output: 
  html_document: 
    toc: yes
    theme: spacelab
    number_sections: yes
    code_folding: hide
---

```{r setup, include=TRUE, message = FALSE, warning = FALSE}
knitr::opts_chunk$set(echo = TRUE)

#attach packages

library(tidyverse)
```

# Confidence Intervals #

In this section, we'll learn to use the 't.test()' function as a way to return the confidene interval using the t-distribution

## Example 1 ##

Create a mock sample of observation values, and find the confidence interval. Then we'll learn how to report that.

```{r}

pika_mass <- c(82, 94, 110, 70, 86, 99, 102, 150)

pika_mean <- mean(pika_mass)

pika_sd <- sd(pika_mass)

pika_ci <- t.test(pika_mass)
```

### Risky & Tedious Way ###

**Writing values out:** 

Mean pika mass at Paiute Pass is 91.9 g $\pm$ 13.5 g (mean $\pm$ standard deviation, n = 7) with a 95% confidence interval of [79.4, 104.4] g.

- Prone to error and have to update constantly if new values given

### High Initial Investment, But Better & Safer Way ###

**In-line code referencing to stored objects:**

Use a single backtick on either side of a lowercase r to create an in-line bit of code

Mean pika mass at Pauite Pass is `r round(pika_mean, 1)` g  $\pm$ `r round(pika_sd, 1)` g (mean $\pm$ standard deviation, n = `r length(pika_mass)`) with a 95% confidence interval [`r round(pika_ci$conf.int[1], 1)`, `r round(pika_ci$conf.int[2], 1)`] g.

# Two-sample t-test #

Use a two-sample t-test to test the null hypothesis that samples were drawn from populations with the same mean (difference in means = 0).

- H0: Means difference = 0
- HA: Means difference is NOT = 0

This is a two-sided t-test because no directionality is implied.

```{r}

oak_trees <- c(29, 19, 22, 30, 35, 16, 20, 7)

pine_trees <- c(48, 32, 41, 40, 56, 70, 30, 19, 20)
```

Is there a significant difference in mean heights for oak and pine trees?

```{r}

trees_t <- t.test(oak_trees, pine_trees)
```

The p-value of `r round(trees_t$p.value, 3)` means that, if these were drawn from populations with the same mean, there is a `r round(trees_t$p.value, 3)*100`% chance of finding sample means *at least as different as those originally found*. This probability is so low that we reject the null hypothesis.



**